## Reproducible Research - Assignment #1
#### Igor Protsenko, Jul 19th, 2014
========================================

```{r echo=FALSE}
library(knitr)
```

### Loading and preprocessing the data

* Loading data from `activity.zip` file

```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")
summary(data)
```

* Processing data into format suitable for analysis


### What is mean total number of steps taken per day?

* Making a histogram of the total number of steps taken each day

```{r}
# aggregating data by date
temp <- aggregate(steps~date,data=data,sum,na.rm=T)
# plotting a histogram
hist(temp$steps, main = "Distribution of Daily Steps", xlab = "Steps")
```

* Calculating **mean** and **median** total number of steps taken per day

```{r}
cat("Mean:",mean(temp$steps),"\n","Median:",median(temp$steps))
```

### What is the average daily activity pattern?

* Making a time series plot of average steps taken per 5-minute interval across days

```{r}
# aggregating data by date
temp <- aggregate(steps~interval,data=data,mean,na.rm=T)
#plotting time series
plot(temp$interval,temp$steps,type="l",
     main = "Average Steps per Interval across Days",
     xlab = "Intervals",
     ylab = "Steps")
```

* Calculating, which 5-minute interval, on average across days, contains maximum number of steps

```{r}
# returning an observation with maximum number of steps
temp[which(temp$steps==max(temp$steps)),]
```

### Imputting missing values

* Calculating total number of rows wih NA's in the dataset

```{r}
# calculating and printing number of rows with NA's
cat("Number of rows with NA's:",length(which(rowSums(is.na(data)) > 0)))
# printing summary of a subset consisiting of NA rows
summary(data[rowSums(is.na(data)) > 0, ])
```

* NA Removal Strategy: substitute NA's with Means for that interval

```{r}
# getting row numbers with NA's
rowsNA <- which(is.na(data$steps))
head(rowsNA)
# calculating Means for each interval across days to use as NA substitute
meansNA <- aggregate(steps~interval,data=data,mean,na.rm=T)
head(meansNA)
```

* Creating a new dataset with NA's filled in with Means for the interval

```{r}
newdata <- data
# substituting NA's for Means for the interval
temp <- data.frame(rowsNA,data[which(is.na(data$steps)),"interval"])
names(temp) <- c("rowsNA","interval")
temp <- merge(temp,meansNA,by="interval")
newdata[rowsNA,"steps"] <- temp[which(temp$rowsNA==rowsNA),"steps"]
# summary of original data with NA's
summary(data)
# summary of new data without NA's (NA's filled in with means for the interval)
summary(newdata)
```

* Making side-by-side histograms to compare two datasets

```{r fig.width=11}
# aggregating data by date
tempOld <- aggregate(steps~date,data=data,sum,na.rm=T)
tempNew <- aggregate(steps~date,data=newdata,sum,na.rm=T)
# plotting two histograms side-by-side
par(mfrow=c(1,2))
hist(tempOld$steps, main = "Daily Steps (data with NA's)", xlab = "Steps")
hist(tempNew$steps, main = "Daily Steps (data without NA's)", xlab = "Steps")
# calculating Mean and Median for old and new datasets
cat("Mean (NA=Yes):",mean(tempOld$steps),"\n","Median (NA=Yes):",median(tempOld$steps),"\n\n",
    "Mean (NA=No):",mean(tempNew$steps),"\n","Median (NA=No):",median(tempNew$steps))
```

### Are there differences in activity patterns between weekdays and weekends?

* Creating a new Weekday/Weekend factor variable

```{r}
# creating new variable
newdata$ww <- weekdays(as.Date(newdata$date))
# changing factor levels to weekday/weekend
newdata[which(
        newdata$ww %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")
              ),"ww"] <- "weekday"
newdata[which(newdata$ww!="weekday"),"ww"] <- "weekend"
newdata$ww <- as.factor(newdata$ww)
summary(newdata)
```

* Creating a panel plot with average steps per interval averaged across weekday/weekend

```{r fig.width=11}
# preparing data for plotting
weekday <- aggregate(steps~interval,data=newdata,mean,subset=newdata$ww=="weekday",na.rm=T)
weekend <- aggregate(steps~interval,data=newdata,mean,subset=newdata$ww=="weekend",na.rm=T)
# plotting two panels side-by-side
par(mfrow=c(1,2))
#plotting "weekday" time series
plot(weekday$interval,weekday$steps,type="l",
     main = "Weekday: Average Steps per Interval",
     xlab = "Intervals",
     ylab = "Steps")
#plotting "weekend" time series
plot(weekend$interval,weekend$steps,type="l",
     main = "Weekend: Average Steps per Interval",
     xlab = "Intervals",
     ylab = "Steps")
```